---
title: "rajee-rf-spread"
author: "rajee"
date: "4/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(curl)
library(tidyverse)
library(readr)
library(dplyr)
library(lubridate)
library(tidyr)
GAMES<-read.csv(url('https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/games.csv'))
GAMES_DETAILS<-read.csv(url('https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/games_details.csv'))
TEAMS<-read.csv(url('https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/teams.csv'))
```

```{r}
GAMES_CLEANING = data.frame(GAMES[,-3])
GAMES = GAMES_CLEANING %>%
  separate(GAME_DATE_EST, sep="-", into = c("GAME_YEAR", "GAME_MONTH", "GAME_DAY"))
```

I'll work on cleaning the TEAMS dataset here
```{r}
TEAMS = data.frame(TEAMS[,-c(1,12)])
TEAMS[3,7] = 17791
TEAMS[13,7] = 19000
TEAMS[15,7] = 20000
TEAMS[17,7] = 19500
TEAMS[18,7] = 18422
```


Taking out rows in Comment column that have characters in them.
```{r}
GAMES_DETAILS<-subset(GAMES_DETAILS, nchar(GAMES_DETAILS$COMMENT)==0)
```

Adding a B(Bench) for empty entries under the Start Position column.
```{r}
GAMES_DETAILS$START_POSITION[GAMES_DETAILS$START_POSITION==""] <- NA
GAMES_DETAILS$START_POSITION[is.na(GAMES_DETAILS$START_POSITION)] = 'B'
```

Changing the minutes in the minute column to seconds.
```{r}
new<- filter(GAMES_DETAILS, nchar(GAMES_DETAILS$MIN)==5)
new1<-filter(GAMES_DETAILS, nchar(GAMES_DETAILS$MIN)==4)
new1$MIN<- paste("0",new1$MIN, sep="")
new2<-filter(GAMES_DETAILS, nchar(GAMES_DETAILS$MIN)==2)
new2$MIN<- paste(new2$MIN, ":00", sep="")
new3<-filter(GAMES_DETAILS, nchar(GAMES_DETAILS$MIN)==1)
new3$MIN<- paste("0", new3$MIN, ":00", sep="")
new4<-filter(GAMES_DETAILS, nchar(GAMES_DETAILS$MIN)==0)
new4$MIN<-paste("00:00", new4$MIN, sep="") 
GAMES_DETAILS<- Reduce(function(...) merge(..., all=TRUE), list(new, new1, new2, new3, new4))
library(lubridate)
mmss_to_ss <- function  (string)
{
  mmss <- strsplit (string, ":", T)
  mm <- as.numeric (mmss[[1]][1])
  ss <- as.numeric (mmss[[1]][2])
  return (mm * 60 + ss)
}
temp<- GAMES_DETAILS[9]
new5<- apply(temp, 1, mmss_to_ss)
new5<- as.data.frame(new5)
GAMES_DETAILS$MIN1 <- new5$new5
GAMES_DETAILS<- subset(GAMES_DETAILS, select = -MIN )
names(GAMES_DETAILS)[28] <- "MIN"
```


Below I am going to calculate eFG for each player in GAMES_DETAILS and then put this variable into our main dataset. To calculate eFG, the formula is (FGM+ (0.5 x 3PM))/FGA. 

```{r}
GAMES_DETAILS <- mutate(GAMES_DETAILS, eFG = (FGM+(0.5*FG3M))/FGA)
head(GAMES_DETAILS)
```


Below is the chunk of code that Mario gave us in class, changed a little bit so that it is for all of the games in the dataset.
```{r}
ALL_GAMES=GAMES %>%
            select(GAME_YEAR, GAME_MONTH, GAME_DAY,GAME_ID,HOME_TEAM_ID,VISITOR_TEAM_ID,PTS_home,PTS_away) %>%
            mutate(Spread=PTS_home-PTS_away,Total=PTS_home+PTS_away)
#Obtain Aggregated OREB from Player Level Statistics
OREB = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,OREB) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(OREB=sum(OREB,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated DREB from Player Level Statistics
DREB = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,DREB) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(DREB=sum(DREB,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated REB from Player Level Statistics
REB = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,REB) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(REB=sum(REB,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated AST from Player Level Statistics
AST = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,AST) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(AST=sum(AST,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated STL from Player Level Statistics
STL = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,STL) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(STL=sum(STL,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated BLK from Player Level Statistics
BLK = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,BLK) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(BLK=sum(BLK,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated TO from Player Level Statistics
TO = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,TO) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(TO=sum(TO,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated FGA from Player Level Statistics
FGA = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,FGA) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(FGA=sum(FGA,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated FTA from Player Level Statistics
FTA = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,FTA) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(FTA=sum(FTA,na.rm=T)) %>%
          ungroup()
#Obtain Aggregated eFG from Player Level Statistics
eFG = GAMES_DETAILS %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,eFG) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(eFG=sum(eFG,na.rm=T)) %>%
          ungroup()
#Starter stats (sum)
Starters_OREB = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, OREB) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(OREB=sum(OREB,na.rm=T)) %>%
  rename(OREB_starters = OREB) %>%
  ungroup()
Starters_DREB = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, DREB) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(DREB=sum(DREB,na.rm=T)) %>%
  rename(DREB_starters = DREB) %>%
  ungroup()
Starters_REB = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, REB) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(REB=sum(REB,na.rm=T)) %>%
  rename(REB_starters = REB) %>%
  ungroup()
Starters_AST = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, AST) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(AST=sum(AST,na.rm=T)) %>%
  rename(AST_starters = AST) %>%
  ungroup()
Starters_STL = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, STL) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(STL=sum(STL,na.rm=T)) %>%
  rename(STL_starters = STL) %>%
  ungroup()
Starters_BLK  = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, BLK) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(BLK=sum(BLK,na.rm=T)) %>%
  rename(BLK_starters = BLK) %>%
  ungroup()
Starters_TO = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, TO) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(TO=sum(TO,na.rm=T)) %>%
  rename(TO_starters = TO) %>%
  ungroup()
Starters_eFG = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, eFG) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(eFG=sum(eFG,na.rm=T)) %>%
  rename(eFG_starters = eFG) %>%
  ungroup()
#Bench stats (sum)
Bench_OREB = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, OREB) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(OREB=sum(OREB,na.rm=T)) %>%
  rename(OREB_bench = OREB) %>%
  ungroup()
Bench_DREB = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, DREB) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(DREB=sum(DREB,na.rm=T)) %>%
  rename(DREB_bench = DREB) %>%
  ungroup()
Bench_REB = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, REB) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(REB=sum(REB,na.rm=T)) %>%
  rename(REB_bench = REB) %>%
  ungroup()
Bench_AST = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, AST) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(AST=sum(AST,na.rm=T)) %>%
  rename(AST_bench = AST) %>%
  ungroup()
Bench_STL = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, STL) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(STL=sum(STL,na.rm=T)) %>%
  rename(STL_bench = STL) %>%
  ungroup()
Bench_TO = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, TO) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(TO=sum(TO,na.rm=T)) %>%
  rename(TO_bench = TO) %>%
  ungroup()
Bench_BLK = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, BLK) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(BLK=sum(BLK,na.rm=T)) %>%
  rename(BLK_bench = BLK) %>%
  ungroup()
Bench_eFG = GAMES_DETAILS %>%
  select(TEAM_ABBREVIATION, GAME_ID, TEAM_ID, eFG) %>%
  filter(GAMES_DETAILS$START_POSITION != 'B') %>%
  group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
  summarize(eFG=sum(eFG,na.rm=T)) %>%
  rename(eFG_bench = eFG) %>%
  ungroup()
```

```{r}
#Merging These Stats Into Game Data
ALL_GAMES_DONE_1 = left_join(ALL_GAMES,select(OREB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(OREB_home=OREB) %>%
                    left_join(select(OREB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(OREB_away=OREB) %>%
                    mutate(OREB=OREB_home+OREB_away)
ALL_GAMES_DONE_2 = left_join(ALL_GAMES_DONE_1,select(DREB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(DREB_home=DREB) %>%
                    left_join(select(DREB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(DREB_away=DREB) %>%
                    mutate(DREB=DREB_home+DREB_away)
ALL_GAMES_DONE_3 = left_join(ALL_GAMES_DONE_2,select(REB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(REB_home=REB) %>%
                    left_join(select(REB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(REB_away=REB) %>%
                    mutate(REB=REB_home+REB_away)
ALL_GAMES_DONE_4 = left_join(ALL_GAMES_DONE_3,select(AST,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(AST_home=AST) %>%
                    left_join(select(AST,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(AST_away=AST) %>%
                    mutate(AST=AST_home+AST_away)
ALL_GAMES_DONE_5 = left_join(ALL_GAMES_DONE_4,select(STL,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(STL_home=STL) %>%
                    left_join(select(STL,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(STL_away=STL) %>%
                    mutate(STL=STL_home+STL_away)
ALL_GAMES_DONE_6 = left_join(ALL_GAMES_DONE_5,select(BLK,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(BLK_home=BLK) %>%
                    left_join(select(BLK,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(BLK_away=BLK) %>%
                    mutate(BLK=BLK_home+BLK_away)
ALL_GAMES_DONE_7 = left_join(ALL_GAMES_DONE_6,select(TO,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(TO_home=TO) %>%
                    left_join(select(TO,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(TO_away=TO) %>%
                    mutate(TO=TO_home+TO_away)
ALL_GAMES_DONE_8 = left_join(ALL_GAMES_DONE_7,select(FGA,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(FGA_home=FGA) %>%
                    left_join(select(FGA,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(FGA_away=FGA) %>%
                    mutate(FGA=FGA_home+FGA_away)
ALL_GAMES_DONE_9 = left_join(ALL_GAMES_DONE_8,select(FTA,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(FTA_home=FTA) %>%
                    left_join(select(FTA,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(FTA_away=FTA) %>%
                    mutate(FTA=FTA_home+FTA_away)
#Merging Starting stats
ALL_GAMES_DONE_10 = left_join(ALL_GAMES_DONE_9,select(Starters_OREB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(OREB_Starters_home=OREB_starters) %>%
                    left_join(select(Starters_OREB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(OREB_Starters_away=OREB_starters) %>%
                    mutate(Starters_OREB=OREB_Starters_home+OREB_Starters_away)
ALL_GAMES_DONE_11 = left_join(ALL_GAMES_DONE_10,select(Starters_DREB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(DREB_Starters_home=DREB_starters) %>%
                    left_join(select(Starters_DREB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(DREB_Starters_away=DREB_starters) %>%
                    mutate(Starters_DREB=DREB_Starters_home+DREB_Starters_away)
ALL_GAMES_DONE_12 = left_join(ALL_GAMES_DONE_11,select(Starters_REB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(REB_Starters_home=REB_starters) %>%
                    left_join(select(Starters_REB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(REB_Starters_away=REB_starters) %>%
                    mutate(Starters_REB=REB_Starters_home+REB_Starters_away)
ALL_GAMES_DONE_13 = left_join(ALL_GAMES_DONE_12,select(Starters_AST,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(AST_Starters_home=AST_starters) %>%
                    left_join(select(Starters_AST,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(AST_Starters_away=AST_starters) %>%
                    mutate(Starters_AST=AST_Starters_home+AST_Starters_away)
ALL_GAMES_DONE_14 = left_join(ALL_GAMES_DONE_13,select(Starters_STL,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(STL_Starters_home=STL_starters) %>%
                    left_join(select(Starters_STL,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(STL_Starters_away=STL_starters) %>%
                    mutate(Starters_STL=STL_Starters_home+STL_Starters_away)
ALL_GAMES_DONE_15 = left_join(ALL_GAMES_DONE_14,select(Starters_BLK,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(BLK_Starters_home=BLK_starters) %>%
                    left_join(select(Starters_BLK,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(BLK_Starters_away=BLK_starters) %>%
                    mutate(Starters_BLK=BLK_Starters_home+BLK_Starters_away)
ALL_GAMES_DONE_16 = left_join(ALL_GAMES_DONE_15,select(Starters_TO,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(TO_Starters_home=TO_starters) %>%
                    left_join(select(Starters_TO,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(TO_Starters_away=TO_starters) %>%
                    mutate(Starters_TO=TO_Starters_home+TO_Starters_away)
#Merging Bench stats
ALL_GAMES_DONE_17 = left_join(ALL_GAMES_DONE_16,select(Bench_OREB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(OREB_Bench_home=OREB_bench) %>%
                    left_join(select(Bench_OREB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(OREB_Bench_away=OREB_bench) %>%
                    mutate(Bench_OREB=OREB_Bench_home+OREB_Bench_away)
ALL_GAMES_DONE_18 = left_join(ALL_GAMES_DONE_17,select(Bench_DREB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(DREB_Bench_home=DREB_bench) %>%
                    left_join(select(Bench_DREB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(DREB_Bench_away=DREB_bench) %>%
                    mutate(Bench_DREB=DREB_Bench_home+DREB_Bench_away)
ALL_GAMES_DONE_19 = left_join(ALL_GAMES_DONE_18,select(Bench_REB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(REB_Bench_home=REB_bench) %>%
                    left_join(select(Bench_REB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(REB_Bench_away=REB_bench) %>%
                    mutate(Bench_REB=REB_Bench_home+REB_Bench_away)
ALL_GAMES_DONE_20 = left_join(ALL_GAMES_DONE_19,select(Bench_AST,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(AST_Bench_home=AST_bench) %>%
                    left_join(select(Bench_AST,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(AST_Bench_away=AST_bench) %>%
                    mutate(Bench_AST=AST_Bench_home+AST_Bench_away)
ALL_GAMES_DONE_21 = left_join(ALL_GAMES_DONE_20,select(Bench_STL,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(STL_Bench_home=STL_bench) %>%
                    left_join(select(Bench_STL,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(STL_Bench_away=STL_bench) %>%
                    mutate(Bench_STL=STL_Bench_home+STL_Bench_away)
ALL_GAMES_DONE_22 = left_join(ALL_GAMES_DONE_21,select(Bench_BLK,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(BLK_Bench_home=BLK_bench) %>%
                    left_join(select(Bench_BLK,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(BLK_Bench_away=BLK_bench) %>%
                    mutate(Bench_BLK=BLK_Bench_home+BLK_Bench_away)
ALL_GAMES_DONE_23 = left_join(ALL_GAMES_DONE_22,select(Bench_TO,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(TO_Bench_home=TO_bench) %>%
                    left_join(select(Bench_TO,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(TO_Bench_away=TO_bench) %>%
                    mutate(Bench_TO=TO_Bench_home+TO_Bench_away)
ALL_GAMES_DONE_24 = left_join(ALL_GAMES_DONE_23,select(eFG,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(eFG_home=eFG) %>%
                    left_join(select(eFG,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(eFG_away=eFG) %>%
                    mutate(eFG=eFG_home+eFG_away)
ALL_GAMES_DONE_25 = left_join(ALL_GAMES_DONE_24,select(Starters_eFG,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(eFG_Starters_home=eFG_starters) %>%
                    left_join(select(Starters_eFG,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(eFG_Starters_away=eFG_starters) %>%
                    mutate(Starters_eFG=eFG_Starters_home+eFG_Starters_away)
ALL_GAMES_DONE_26 = left_join(ALL_GAMES_DONE_25,select(Bench_eFG,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(eFG_Bench_home=eFG_bench) %>%
                    left_join(select(Bench_eFG,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(eFG_Bench_away=eFG_bench) %>%
                    mutate(Bench_eFG=eFG_Bench_home+eFG_Bench_away)
#Creating Home Team and Away Team Variables
ALL_TEAMS= TEAMS %>%
            select(TEAM_ID,CITY,NICKNAME) %>%
            unite(NAME,CITY,NICKNAME,sep=" ")
#Merging Team Name into original data
ALL_GAMES_DONE_AGAIN=left_join(ALL_GAMES_DONE_26,ALL_TEAMS,by=c("HOME_TEAM_ID"="TEAM_ID")) %>%
                        rename("Home Team"=NAME) %>%
                        left_join(ALL_TEAMS,by=c("VISITOR_TEAM_ID"="TEAM_ID")) %>%
                        rename("Away Team"=NAME) %>%
                        select("Home Team","Away Team",everything()) %>%
                        select(-GAME_ID,-HOME_TEAM_ID,-VISITOR_TEAM_ID)
head(ALL_GAMES_DONE_AGAIN)
```

```{r rf-model-for-spread}
library(randomForest)
data2014 = subset(ALL_GAMES_DONE_AGAIN, GAME_YEAR == 2014)
data2015 = subset(ALL_GAMES_DONE_AGAIN, GAME_YEAR == 2015)
data2016 = subset(ALL_GAMES_DONE_AGAIN, GAME_YEAR == 2016)
data2017 = subset(ALL_GAMES_DONE_AGAIN, GAME_YEAR == 2017)
data2018 = subset(ALL_GAMES_DONE_AGAIN, GAME_YEAR == 2018)
data2019 = subset(ALL_GAMES_DONE_AGAIN, GAME_YEAR == 2019)
data2020 = subset(ALL_GAMES_DONE_AGAIN, GAME_YEAR == 2020)

final2014 <- data2014[,-c(1,2,3,4,5,6,7,9,79,80,81,82,83,84,85,86,87)]
final2015 <- data2015[,-c(1,2,3,4,5,6,7,9,79,80,81,82,83,84,85,86,87)]
final2016 <- data2016[,-c(1,2,3,4,5,6,7,9,79,80,81,82,83,84,85,86,87)]
final2017 <- data2017[,-c(1,2,3,4,5,6,7,9,79,80,81,82,83,84,85,86,87)]
final2018 <- data2018[,-c(1,2,3,4,5,6,7,9,79,80,81,82,83,84,85,86,87)]
final2019 <- data2019[,-c(1,2,3,4,5,6,7,9,79,80,81,82,83,84,85,86,87)]

testing <- data2020[,-c(1,2,3,4,5,6,7,9,79,80,81,82,83,84,85,86,87)]
updated <- subset(ALL_GAMES_DONE_AGAIN, GAME_YEAR>=2014)
updated2 <- subset(updated, GAME_YEAR<2020)
training <- updated2[,-c(1,2,3,4,5,6,7,9,79,80,81,82,83,84,85,86,87)]

head(training)
head(testing, n=50)

rf_spread <- randomForest(Spread ~ ., data = training, importance = TRUE,proximity=TRUE,oob.prox=TRUE)
rf_spread

predictions_train <- predict(rf_spread, training)
predictions_train

predictions_test <- predict(rf_spread,testing)
predictions_test

```

```{r testing}
# install.packages("caret")
# library(caret)
# 
# folds <- 10
# cvIndex <- createFolds(factor(ALL_GAMES_DONE_AGAIN$Spread), folds, returnTrain = T)
# tc <- trainControl(index = cvIndex,
#                    method = 'cv', 
#                    number = folds)
# 
# 
# preProcValues <- preProcess(ALL_GAMES_DONE_AGAIN, method = c("center","scale"))
# train_processed <- predict(preProcValues, ALL_GAMES_DONE_AGAIN)
# train_processed$Spread <- ifelse(train_processed$Spread==FALSE,0,1)
# 
# index <- createDataPartition(train_processed$GAME_YEAR, p=0.75, list=FALSE)
# trainSet <- train_processed[ index,]
# testSet <- train_processed[-index,]
# 
# control <- rfeControl(functions = rfFuncs,
#                       method = "repeatedcv",
#                       repeats = 3,
#                       verbose = FALSE)
# 
# outcomeName<-'Spread'
# predictors<-names(trainSet)[!(names(trainSet) %in% outcomeName | names(trainSet) == "Spread")]
# #install.packages("e1071")
# complete.cases(trainSet)
# any(is.na(trainSet))
# Quote_Pred_Profile <- rfe(trainSet[,predictors], (trainSet[,outcomeName]),
#                           na.action=na.omit,
#                           metric = "Accuracy",
#                           maximizize = TRUE,
#                          rfeControl = control, na.omit=TRUE)
# plot(Quote_Pred_Profile)
# summary(Quote_Pred_Profile)
```

